On Cybersecurity\\Science and Engineering\footnote{Confidential, restricted to the authors. Property of V-Research S.r.l.}

### 2020-04-07 formal_glossary.tex
review of Theory section

The main outcome of this section is that there is no evidence that properly
considering the malicious intent of an agent necessarily results in a security
scientific theory.  On the other hand, the complete definition of Errors,
predicted by a (sound w.r.t. Errors) security theory seems to be a proper
research direction.  In other words, system \emph{in-security} and
\emph{security} should not be searched on those events who maliciously produces
negative effects on a systems (incidents), but in the potential designs and
implementations due to vague specifications. Therefore, we will consider a
security-by-design approach based on a requirement-based engineering process.

>>>>

The main outcome of this section is that there is no evidence that properly
considering the malicious intent of an agent necessarily results in a falsifiable or general security
theory.  On the other hand, the complete definition of Errors,
predicted by a (sound w.r.t. Errors) security theory seems to be a proper
research goal.  In other words, system \emph{in-security} and
\emph{security} should not be searched on those malicious events which produces
negative effects on a systems (incidents), but in the potential designs and
implementations due to vague specifications. Therefore, we will consider a
security-by-design approach based on a requirement-based engineering process.

### 2020-04-07 theory.tex
review of Theory section

In order to address the problem raised by Herley, 
we shall define how to distinguish between a secure and an insecure system.
While most of the literature (and this paper so far) have
correlated the problem of in-security to the maliciousness of
agents interacting with the system, we show that security
doesn't seem to stem from a malicious nature but, rather, insecurity 
raises from the lack of well-defined security requirements for the
design process of a system. We argue that the high number
of security vulnerability reported today are simply the realization
of potential system configurations, which deviate from the nominal
behavior just because the \emph{intended behavior}\fixnote{mr}{maybe ``nominal'' instead of ``intended'' is better and behavior can be confused with the
relation between assertions and beliefs which is not what we mean} of the system is not precisely 
defined in the specification at the very early stages of the engineering process. 
A reference engineering process (so called, system security lifecycle of a product)
is described, in details, in Section~\ref{sec:process}. However,
for the sake of simplicity, the reader can assume for this section that 
we are currently dealing with just the relation between 
the specification of a system and its design.

@@

\subsection{A Reference Model for Cybersecurity Engineering}\label{sec:vmodel}
In order to describe the $\abf$-theory we briefly introduce the 
engineering concepts (detailed and extended in Section~\ref{sec:engineering})
underlying the $\abf$-theory; namely, the first three stages of the, so called,
engineering V-Model. Our definition is partial and we only use it for the 
sake of simplicity and to better present the theory.

The development process of a new product (a CPS in our case) 
should follow an engineering model. Our reference engineering process 
(inspired by the waterfall V-Model\autocite{waterfallvmodel}), as depicted in
Figure~\ref{fig:simplifiedvmodel}, is structured as the following ordered list
of abstraction refinement steps.
\begin{enumerate}
	\item The process starts with the definition of a \emph{specification}
		where the general requirements of the CPS are defined (e.g.
		functional, physical, network, and design).  As an example, an
		RFC of a communication or security protocol can be considered
		as the result of this first phase (e.g. see
		RFC-1\autocite{rfc1}.  In this first phase, we consider the
		architectural aspects of a CPS such as the draft physical and
		functional architectures.
	\item The process then continues with the definition of a design where
		the requirements of the previous phase holds; detailing the
		logic of the physical (e.g. the physical-layer encoding of the
		information) and functional architectures (e.g. the protocol
		logic of communication over a number of potential sessions).
	\item Finally, the design is implemented as hardware (HW) for the physical architecture
		and software (SW) for the functional architecture.
\end{enumerate}

\subsection{Abstract System}\label{sec:systemstate}
The ISO/IEC/IEEE 15288:2015 (System Life Cycle Processes) provides a definition
of system as ``A combination of interacting elements organized to achieve one
or more stated purposes.''\autocite{ISO201515288}.  Therefore, a system can be
considered as a single agent where its interacting elements are the
constituents of the agent itself. For the sake of simplicity we first define a
system as an agent and then extend the definition to a ``combination of
interacting'' agents.  We use the same process of \autocite{Santaca2016abf} but
over slightly different terms. The differences and commonalities with
\autocite{Santaca2016abf} are described in Appendix~\ref{app:abf}.

There is no agreement between the research communities (e.g.
Multi-Agent-System, Epistemic Logic) on which are the constituent of an agent
as a system. However, the same ideas revolved around for thousands of years.
Some relevant examples for our objective are the following.
\begin{itemize}
	\item In \autocite{Hintikka1962knowledge}, Hintikka describes the
		difference between Knowledge and Belief (as epistemological
		concepts), and the whole Doxastic logic defines in details how
		Beliefs can be formalized.
	\item In \autocite{Hintikka1993Information}, Hintikka describes the concept
		of Information and the difference with Knowledge and Belief.
	\item In \autocite{Empiricus1990Pyrrhonism}, the author states: ``The
		logical criterion also may be used in three senses -- of the
		agent, or the instrument, or the ``according to what''; the
		agent, for instance, may be a man, the instrument either sense
		perception or intelligence, and the ``according to what'' the
		application of the impression ``according to'' which the man
		proceeds to judge by means of one of the aforesaid
		instruments.'' 
	\item In \autocite{Santaca2016abf}, the authors defines an agent as a
		tuple of Assertions, Beliefs, and Facts.
\end{itemize}

If we consider \autocite{Empiricus1990Pyrrhonism}, the author applies a
``logical criterion'' to an agent, for instance a man, for instance a device in
a system (or a system) as in \autocite{Santaca2016abf}.  The instrument identified is sense
perception, because a man is considered as example, but we can abstract from the
concept of instrument and consider any received information, where information is
considered as in \autocite{Hintikka1993Information}; which can be
summarized by the following axiom: ``Information is specified by specifying
which alternatives concerning the reality it admits and which alternatives
excludes''. This means, for our argument, that considering a propositional
variable, which admits the two alternatives True/False, its information is
defined as Believed to be True/False and not Believed to be the opposite. 
%Similarly, when reasoning on a
%mereotopological region w.r.t. the RCC (e.g. RCC3 admits two regions to be
%equal, disjoint, or overlapping) information on this region should exclude
%some of the alternatives. 
Finally, in \autocite{Empiricus1990Pyrrhonism}
the author states that the logical criterium can be judged according to a frame of reference.
For our argument, as in \autocite{Santaca2016abf}, we define this concept
as Knowledge (or Facts); where knowledge is defined, as in \autocite{Steup2020epistemology}, as
a set of proposition known by an agent, such that: (i) knowledge requires belief,
(ii) knowledge require truth, (iii) knowledge must be \emph{properly justified}.

Before detailing what ``properly justified'' means, we discuss the mentioned
relation between agent and belief and we introduce a more formal definition
of knowledge and belief. For our argument, the only objective of
Information is to exchange beliefs between agents.
Due to the definition of Information and then with its relation to the probabilistic
correlation to truth/reality, we consider
information in relation with an agent's beliefs. Similarly, we consider
beliefs to define the actual behavior of an agent or a system. On the other hand,
Knowledge drives the nominal behavior of an agent or a system. This view
is not considered in most (if not all) the approaches to protocol verification
and the Dolev-Yao theory is usually applied to a representation/abstraction
of a protocol that only considers knowledge and transfer of knowledge (i.e.
if an agent sends a message, the recipient knows that message). However,
the security of a system is tightly related to the difference between
the nominal behavior and the actual (e.g. the implemented) one.

\begin{definition}{\bf System State and State-Space --}\label{def:system}
	The state of a system (or a sub-system) $s$ is identified with an
	instance of an agent state (see Definition~\ref{def:agent}) such that
	$s=\langle\rcc(\knowledgeRegion,\beliefRegion),\rcc(\knowledgeRegion,\informationRegion),\rcc(\beliefRegion,\informationRegion)\rangle$,
	where $\knowledgeRegion,\beliefRegion,\informationRegion$ are Regions
	of Knowledge, Beliefs, and Information respectively, and $\rcc$ is a
	specific relation between the Regions it predicates on.  The
	state-space of a system is then defined by the different $\rcc$
	relations between the three pairs of Regions defining the system and
	all the sub-systems.
\end{definition}
Therefore, a system state space can be seen as a superposition of multiple agent states where knowledge,
beliefs, and information are related between each other (in a mereotopological
space) through all the admissible RCC relations. 
\fixnote{mr}{are the next lines redundant?}A \emph{system state} (as an instance of the system state space)
is a specific configuration of
the $\rcc$ relations between the three regions. Therefore, a collection of 
system states defines a collection of choices for those $\rcc$ relations.

@@

\begin{definition}{\bf Knowledge --}\label{def:knowledge}
Given an abstract collection of Agents $Ag$, and the modal operator
	$\knows{a}{}$ (where $a\in Ag$), Knowledge is defined as a region 
	of predicates known by an agent $\knowledge{a}=\bigcup_\Phi \knows{a}{\varphi}$ 
	(where $\Phi$ is the collection of all the propositions known by $a$).
	Given a proposition $P$, we extend the semantics of the Causality structure with:
	\begin{enumerate}[noitemsep]
		\item[$(\interpretation16)$] $\world\models\knows{a}{P}$ iff
			$\world'\models P$ for all $\world'$ such that
			$\world\modalrelation\world'$.
	\end{enumerate}
\end{definition}

@@

\begin{enumerate}
	\item We consider Information only when the intention of exchanging 
		that Information is from a sender to 
		recipient is defined; and we call it \emph{Assertion}  
	\item Similarly, the portion of Beliefs we consider for system
		engineering is the one that builds (input) or describes
		(output) the behavior of an agent (strategy rules
		\footnote{``The logical structure of information is one of the
		most basic and one of the most basic and one of the simplest
		thing in the wide and wonderful world of logical analysis. This
		point can be put in a deeper perspective. A distinction
		[\ldots] ought to be made [--] between two kinds of rules (or
		principles) in any strategic activity like knowledge seeking.
		On the one hand you have the rules that define the game, e.g.
		how chessmen are moved on a board. The can be called
		\emph{definitory} rules.  They must be distinguished from rules
		[\ldots] that deal with what is better and what is worse in the
		game in question.  Definitory rules do not say anything about
		this subject. Rules which do can be called \emph{strategic
		rules}'' -- Hintikka in \autocite{Hintikka1993Information}}),
		and
	\item We consider a set of axiomatic \emph{Facts} (definitory rules\fixnote{mr}{facts are definitory rule, as they don't define how a real system is finally implemented (curently) but how it should, trugh a series of requirements which may not be respected, through insecurity, in the implemented system}) instead
		of considering the more general epistemic definition of
		Knowledge. Specifically, Facts describes:
		\begin{itemize}
			\item The functional architecture of each agents
			\item The physical/structural (HW/SW) architecture of each
				subsystem of agents and agent within a
				subsystem
			\item Other important aspects of the system that we
				describe afterwards in this paper, such as
				the assets and security properties 
		\end{itemize}
\end{enumerate}
We give a graphical representation of (sub-)system and agent in Figure~\ref{fig:system-agent}.

@@

\begin{definition}{\bf Assertion -- }\label{def:assertion}
	An assertion (\fix{mr}{subtype of Information}) of  is an intended transfer of beliefs between
	two agents $a$ and $b$ such that
	\begin{enumerate}[noitemsep]
		\item[$(\interpretation19)$] $\world\models\rassert{a}{b}{P}$ iff
			$\world\models\belief{a}{P}$, 
			$\world\models\neg\belief{a}{\neg P}$, 
			$\world'\models\rcc(\belief{a}{P},\belief{b}{P})$, 
			for all $\world'$ such that $\world\modalrelation\world'$.
	\end{enumerate}
\end{definition}

\begin{definition}{\bf Behavior -- }\label{def:behavior}
	The behavior (\fix{mr}{subtype of Beliefs}) of an agent is defined as the transformation process (e.g. defined as a
	protocol, or a functional architecture) that determines the
	output-beliefs based on input-beliefs and vice versa (where input-beliefs and
	output-beliefs are beliefs taken as input or output respectively).
	\begin{enumerate}[noitemsep]
		\item[$(\interpretation20)$] $\world\models\bigcup_{\Phi}\believe{s}{\varphi}$ and $\world\models\behave{a}{[\bigcup_{s\in S}\believe{s}{P_{s}}]}$ iff
			$\world\models\belief{a}{P'}$,

		where $a$ is an agent, $S$ is a set of agents such that $a\neg\in S$, and the calculation of $P'$ is deteministic.
	\end{enumerate}
\end{definition}

\begin{definition}{\bf Fact -- }\label{def:fact}
	Facts are a \fix{mr}{sub-region/subtype} of Knowledge $\factRegion = \bigcup_\Phi \knows{a}{\varphi}$
	that predicate in a factive way, e.g.
	if $a$ knows that $P$ then $P$, and the region of facts is monotone (no revision).
	\begin{enumerate}[noitemsep]
		\item[$(\interpretation21)$] $\world\models\fact{a}{P}$ iff
			$\world \models \Box P$. 
	\end{enumerate}
\end{definition}

\begin{definition}{\bf Operational System State --}\label{def:system}
	An operational system (or a sub-system) is represented as an agent (see Definition~\ref{def:agent}) and then defined by the resulting state space,
	such that
	$s=\langle\rcc(\factRegion,\behaviorRegion),\rcc(\factRegion,\assertionRegion),\rcc(\behaviorRegion,\assertionRegion)\rangle$,
	where $\assertionRegion,\behaviorRegion,\factRegion$, are regions of assertions, behavior (i.e. the beliefs generated by the behavior), and facts respectively.
\end{definition}

@@

In the quantitative evaluation of a single agent, depicted in Figure~\ref{fig:quantitative},
we argue that only 1 configuration represents the nominal (expected) behavior 
of the agent while the other configurations are either impossible to 
implement or diverge from the intended nominal behavior. We note 
that, the numbers reported here do not consider the details of the
engineering process and should be considered a limit of an abstract 
representation of the system. In Section~\ref{sec:engineering} we
detail this numeric evaluation to faitfully represent the
number of insecurity configurations.

@@

As described beforehand in this section, the collection of Facts shall define
the functional and the physical architecture of an agent, but Assets and
security properties are defined as Facts, i.e. true. Given that understanding
why an Asset is considered as such by the user doesn't necessarily allow
us to better qualify an Asset, we consider the fact of ``being an asset'' as
a predicate \fix{mr}{(or a true function)} of a component of a system.

@@

\subsection{Qualitative Evaluation of Agent Space in $\abf$}\label{sec:agentspace}
While a quantitative analysis reveals how many possible configurations of an
agent (i.e. a system) exist w.r.t. the $\abf$ theory (i.e. $54/125$ in RCC5), a
qualitative analysis of the different configurations describes
the configurations allowed by the $\abf$ theory, and how those configurations can be
categorized.  In Table~\ref{tab:5com} we provide the generic composition table
of RCC5 over 3 regions instantiated over $\abf$, which shows the whole state
space for a single agent. The color coding of the table represents the 
risk function as a risk matrix as we detail afterwards in this section.

A detailed qualitative categorization of agents over a theory similar to the
$\abf$ one, is presented in this article and described in
\autocite{Santaca2016abf}. The differences in the formalization and of the
application requires another in-depth analysis. As depicted in
Figure~\ref{fig:soundness}, the relation between Facts, and Assertions and Behavior
defines the soundness of the design (admissible configurations of Assertions,
Behaviors and, in turn, Beliefs) w.r.t. the specification (what the
specification mandates, such as by the nature of the physical/functional
architectural models). We categorize agents by first analyzing the relations
between each pair of Regions defining an agent (i.e. $\abf$), and then we
categorize the different agents as tuple of the three Regions.
For the sake of simplicity, soundness is opposed to non-soundness in the following, however,
with the RCC we consider different ``degrees'' of soundness. For example,
in RCC5, if we consider EQ between two Regions as representing soundness, DR over the same Regions
represents non-soundness while PP, PO, PPi represents the different degrees of soundness.

@@

\Paragraph{\Rcc{$\assertionRegion$}{$\behaviorRegion$} -- Collaboration}
In order to reason on the relation between assertions and behavior we first
need to consider that, by definition, assertions are defined as transfer of
information between two agents, e.g., $a$ and $b$.  Therefore, as depicted in
Figure~\ref{fig:system-agent}, an agent has two main categories of assertions,
input and output assertions.  Given an agent $a$ and a collection of asserted
predicates $\Phi$, the Input assertions are those received by $a$ from an agent
$s$ acting as a sender, $\rassert{s}{a}{\Phi}$; similarly, output assertions
are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. We shall consider
two pairs of regions\footnote{``I am not asserting, as Lotze did, that a relation between $X$ and $Y$
consists of a quality in $X$ and a quality in $Y$ -- a view which I regard as
quite indefensible. -- I assert that a relation $Z$ between $X$ and $Y$ involves
the existence in $X$ of the quality ``having the relation $Z$ to $Y$'' so that
a difference of relations always involves a difference in quality, and a
change of relations always involves a change of quality.'' --  Ellis J. McTaggart, The Unreality of Time\autocite{Mctaggart1908unreality}}: 
\begin{itemize}
	\item \Rcc{$\assertionRegion_{s\rightarrow a}$}{$\behaviorRegion$},
		where the relation between Input-assertions and behavior
		describes the soundness of the execution of the functional
		architecture w.r.t. input elicitation. With more details, the
		ideal specification of the functional architecture, along with
		the expected inputs, defines the functional behavior of an
		ideal system.  If all the inputs (Assertions) are correctly
		handled in the functional specification (behavior) the
		specification is sound.

@@

\Paragraph{\Rcc{$\assertionRegion$}{$\factRegion$} and \Rcc{$\behaviorRegion$}{$\factRegion$} -- Honesty and Competence}
While Assertions and Beliefs generated by the Behavior determines how a system should work, and the
relation between the two defines the quality (i.e. the completeness w.r.t the specification) of the agent, the relation of
Assertions and Behavior with Facts determines the quality with respect to the
nominal (specified) system.  Given that Facts defines what is true in the
system (e.g. the level of water in a tank may be considered as a Fact, while
the reading of a sensor as a Belief), the relation of Assertions an Facts
determines the degree of quality between the information circulating in a
system (or within an agent) and reality\fixnote{mr}{assertions are reality
while facts are dogmatic assertions at specification level, i.e. requirements}.
Since the transfer of information through Assertions
generates Beliefs, a dishonest agent may circulate false information,
generating false Beliefs.  We note that it is often implied that the intention behind
circulating false information discerns a dishonest and an incompetent agent,
however, we consider honesty related to truth rather than related to what is
true\footnote{``[\ldots] truth exists only in the good man, but the true in the bad
man as well; for it is possible for the bad man to utter something true.''
Sextus Empiricus, Outlines of Pyrrhonism, II-83\autocite{Empiricus1990Pyrrhonism}}.  This implies that the
relation between Beliefs and Facts determines the competence (on the subjected
defined by the Facts) of an agent (i.e. the more competent an agent is, the
more likely a Belief of that agent is true).

>>>>

In order to address the problem raised by Herley, we shall define how to
distinguish between a secure and an insecure system.  While most of the
literature have correlated the problem of in-security
to the maliciousness of agents interacting with the system, we show that
security doesn't seem to stem from a malicious nature but, rather, insecurity
raises from the lack of well-defined security requirements for the design
process of a system. We argue that the high number of security vulnerability
reported today are simply the realization of potential system configurations,
which deviate from the nominal behavior just because the \emph{intended (or
nominal) behavior} of the system is not precisely defined in the specification
at the very early stages of the engineering process.

\subsection{A Reference Model for Cybersecurity Engineering}\label{sec:vmodel}
A reference engineering process (so called, system security lifecycle of a
product) is described, in Section~\ref{sec:engineering}.  For the sake of
simplicity, we now give an high-level overview of a reference engineering model
focusing on the first three stages of the, so called, engineering V-Model. Our
definition is partial and we only use it for the sake of simplicity and to
better present the theory.

The development process of a new product (a CPS in our case) 
should follow an engineering model. Our reference engineering process 
(inspired by the waterfall V-Model), as depicted in
Figure~\ref{fig:simplifiedvmodel}, is structured as the following ordered list
of abstraction refinement steps.
\begin{enumerate}
	\item The process starts with the definition of a \emph{specification}
		where the general requirements of the CPS are defined (e.g.
		functional, physical, network, and design).  As an example, an
		RFC of a communication or security protocol can be considered
		as the result of this first phase (e.g. see
		RFC-1\autocite{rfc1}.  In this first phase, we consider the
		architectural aspects of a CPS such as the draft physical and
		functional architectures.
	\item The process then continues with the definition of a design where
		the requirements of the previous phase holds; structuring the
		physical architecture along with its logic (e.g. the physical-layer encoding of the
		information), and the functional architectures and its logic (e.g. the protocol
		logic of communication over a number of potential sessions).
	\item Finally, the design is implemented as hardware (HW) for the physical architecture
		and software (SW) for the functional architecture.
\end{enumerate}

@@

\subsection{Abstract System}\label{sec:systemstate}
The ISO/IEC/IEEE 15288:2015 (System Life Cycle Processes) provides a definition
of system as ``A combination of interacting elements organized to achieve one
or more stated purposes.''\autocite{ISO201515288}.  Therefore, a system can be
considered as a single agent where its interacting elements are the
constituents of the agent itself. For the sake of simplicity we first define a
system as an agent and then extend the definition to a ``combination of
interacting'' agents.

There is no agreement between the research communities (e.g.
Multi-Agent-System, Epistemic Logic) on which are the constituent of an agent
as a system. However, the same ideas revolved around for thousands of years.
Some relevant examples for our objective are the following.
\begin{itemize}
	\item In \autocite{Hintikka1962knowledge}, Hintikka describes the
		difference between Knowledge and Belief (as epistemological
		concepts), and the whole Doxastic logic defines in details how
		Beliefs can be formalized.
	\item In \autocite{Hintikka1993Information}, Hintikka describes the concept
		of Information and the difference with Knowledge and Belief.
	%\item In \autocite{Empiricus1990Pyrrhonism}, the author states: ``The
	%	logical criterion also may be used in three senses -- of the
	%	agent, or the instrument, or the ``according to what''; the
	%	agent, for instance, may be a man, the instrument either sense
	%	perception or intelligence, and the ``according to what'' the
	%	application of the impression ``according to'' which the man
	%	proceeds to judge by means of one of the aforesaid
	%	instruments.'' 
	\item In \autocite{Santaca2016abf}, the authors defines an agent as a
		tuple of Assertions, Beliefs, and Facts.
\end{itemize}

\Paragraph{Methodology}
Before going into the details of the various terms and their applications for
the definition of our theory, we 
detail our methodology.
The concept of Knowledge \autocite{Hintikka1962knowledge} (from an
epistemological point of view) can be defined as the way to get to the Truth.
Anything that is known is considered true\footnote{It is also
believed that Knowledge, defined as we did, is in-apprehensible \autocite{Empiricus1990Pyrrhonism}, but this shouldn't be an issue since it is a common problem to every scientific search}
In contrast, Beliefs \autocite{Hintikka1962knowledge} are potentially true and
leads to new hypothesis. Due to the
structure of deductive reasoning we can only start from unproven hypothesis and test the
deductive conclusions on the reality. Therefore, whatever a scientific-security-theory may be,
it can only be defined starting from hypotheses.  A sound reasoning should lead to
conclusion that shall be tested on a real system. Therefore, only at the end of
the reasoning we will be able to test, and judge, the quality of the theory
itself. The identification of hypotheses should start from probably true
beliefs (e.g. from Information \autocite{Hintikka1993Information}), and then from the state of the art.

For our argument, 
as in \autocite{Santaca2016abf}, an agent is composed by its knowledge, beliefs, and the information or assertions
it provides; where knowledge is defined, as in \autocite{Steup2020epistemology}, as
a set of proposition known by an agent, such that: (i) knowledge requires belief,
(ii) knowledge require truth, (iii) knowledge must be \emph{properly justified}, and
the only objective of Information is to exchange beliefs between agents.
As defined in \autocite{Hintikka1993Information}, ``Information is specified by specifying
which alternatives concerning the reality it admits and which alternatives
excludes''. This means that if we consider a propositional
variable (which admits the two alternatives True/False) its information is
defined as Believed to be True/False and not Believed to be the opposite. 
Due to the definition of Information and then with its relation to the probabilistic
correlation to truth/reality, we consider
information in relation with an agent's beliefs. Similarly, we consider
beliefs to define the actual behavior of an agent or a system. On the 
contrary, this view is not considered in most (if not all) the approaches to formal protocol verification
in the symbolic model. In fact, the security reasoning is usually applied to a representation/abstraction
of a protocol that only considers knowledge and transfer of knowledge (i.e.
if an agent sends a message, the recipient knows that message). However,
we believe that the security of a system is tightly related to the differences between
the nominal behavior and the actual (e.g. the implemented) one and only in the
nominal one (as hypothetical) one should consider agent's knowledge.

\begin{definition}{\bf System State and State-Space --}\label{def:system}
	The state of a system (or a sub-system or an agent) $s$ is defined as the tuple 
	$s=\langle\rcc(\knowledgeRegion,\beliefRegion),\rcc(\knowledgeRegion,\informationRegion),\rcc(\beliefRegion,\informationRegion)\rangle$,
	where $\knowledgeRegion,\beliefRegion,\informationRegion$ are Regions
	of Knowledge, Beliefs, and Information respectively, and $\rcc$ is a
	specific relation between Regions.  The
	state-space of a system is then defined by the different $\rcc$
	relations between the three pairs of Regions defining the system and
	all the sub-systems.
\end{definition}
Therefore, a system state space can be seen as a superposition of multiple agent states where knowledge,
beliefs, and information are related between each other (in a mereotopological
space) through all the admissible RCC relations. 
A \emph{system state} (as an instance of the system state space)
is a specific configuration of
the $\rcc$ relations between the three regions. Therefore, a collection of 
system states defines a collection of choices for those $\rcc$ relations.

@@

\begin{definition}{\bf Knowledge --}\label{def:knowledge}
Given an abstract collection of Agents $Ag$, and the modal operator
	$\knows{a}{}$ (where $a\in Ag$), Knowledge is defined as a region 
	of predicates known by an agent $\knowledge{a}=\bigcup_\Phi \knows{a}{\varphi}$ 
	(for all $\varphi\in\Phi$ where $\Phi$ is the collection of all the propositions known by $a$).
	Given a proposition $P$, we extend the semantics of the Causality structure with:
	\begin{enumerate}[noitemsep]
		\item[$(\interpretation16)$] $\world\models\knows{a}{P}$ iff
			$\world'\models P$ for all $\world'$ such that
			$\world\modalrelation\world'$.
	\end{enumerate}
\end{definition}

@@

\begin{enumerate}
	\item We consider Information only when the intention of exchanging 
		that Information is from a sender to 
		recipient is defined; and we call it \emph{Assertion}  
	\item Similarly, the portion of Beliefs we consider for system
		engineering is the one that builds (input) or describes
		(output) the behavior of an agent (strategy rules
		\footnote{``The logical structure of information is one of the
		most basic and one of the most basic and one of the simplest
		thing in the wide and wonderful world of logical analysis. This
		point can be put in a deeper perspective. A distinction
		[\ldots] ought to be made [--] between two kinds of rules (or
		principles) in any strategic activity like knowledge seeking.
		On the one hand you have the rules that define the game, e.g.
		how chessmen are moved on a board. The can be called
		\emph{definitory} rules.  They must be distinguished from rules
		[\ldots] that deal with what is better and what is worse in the
		game in question.  Definitory rules do not say anything about
		this subject. Rules which do can be called \emph{strategic
		rules}'' -- Hintikka in \autocite{Hintikka1993Information}}),
		and
	\item We consider a set of axiomatic \emph{Facts} (definitory rules\fixnote{mr}{facts are definitory rule, as they don't define how a real system is finally implemented (curently) but how it should, trugh a series of requirements which may not be respected, through insecurity, in the implemented system}) instead
		of considering the more general epistemic definition of
		Knowledge. Specifically, Facts describes:
		\begin{itemize}
			\item The functional architecture of each agents
			\item The physical/structural (HW/SW) architecture of each
				subsystem of agents and agent within a
				subsystem
			\item Other aspects of the system, related to the two previous points, that we
				describe afterwards in this paper, such as
				the assets and security properties 
		\end{itemize}
\end{enumerate}
We give a graphical representation of (sub-)system and agent in Figure~\ref{fig:system-agent}.

@@

\begin{definition}{\bf Assertion -- }\label{def:assertion}
	An assertion is an intended transfer of beliefs between
	two agents $a$ and $b$ such that
	\begin{enumerate}[noitemsep]
		\item[$(\interpretation19)$] $\world\models\rassert{a}{b}{P}$ iff
			$\world\models\belief{a}{P}$, 
			$\world\models\neg\belief{a}{\neg P}$, 
			$\world'\models\rcc(\belief{a}{P},\belief{b}{P})$, 
			for all $\world'$ such that $\world\modalrelation\world'$.
	\end{enumerate}
\end{definition}

\begin{definition}{\bf Behavior -- }\label{def:behavior}
	The behavior of an agent is defined as the transformation process (e.g. defined as a
	protocol, or a functional architecture) that determines the
	output-beliefs based on input-beliefs and vice versa (where input-beliefs and
	output-beliefs are beliefs taken as input or output respectively).
	\begin{enumerate}[noitemsep]
		\item[$(\interpretation20)$] $\world\models\bigcup\behaviorRegion(\beliefRegion_I,\beliefRegion_O)$ and 
			$\world\models\beliefRegion_I\wedge\beliefRegion_O$,
		where $a$ is an agent, $\beliefRegion_I$ and $\beliefRegion_O$ are input and output-beliefs respectively.
	\end{enumerate}
\end{definition}

\begin{definition}{\bf Fact -- }\label{def:fact}
	Facts are defined as $\factRegion = \bigcup_\Phi \knows{a}{\varphi}$
	that predicate in a factive way, e.g.
	if $a$ knows that $P$ then $P$, and the region of facts is monotone (no revision).
	\begin{enumerate}[noitemsep]
		\item[$(\interpretation21)$] $\world\models\fact{a}{P}$ iff
			$\world \models \Box P$. 
	\end{enumerate}
\end{definition}

\begin{definition}{\bf Operational System State --}\label{def:opsystem}
	An operational system (or a sub-system) state is defined as the tuple
	$s=\langle\rcc(\factRegion,\behaviorRegion),\rcc(\factRegion,\assertionRegion),\rcc(\behaviorRegion,\assertionRegion)\rangle$,
	where $\assertionRegion,\behaviorRegion,\factRegion$, are regions of assertions, behavior (i.e. the beliefs generated by the behavior), and facts respectively.
\end{definition}

@@

In the quantitative evaluation of a single agent, depicted in Figure~\ref{fig:quantitative},
we argue that only 1 configuration represents the nominal (expected) behavior 
of the agent while the other configurations are either impossible to 
implement or diverge from the intended nominal behavior. We note 
that, the numbers reported here do not consider the details of the
engineering process and should be considered a limit of an abstract 
representation of the system. In Section~\ref{sec:engineering} we
detail this numeric evaluation to faithfully represent the
number of insecurity configurations, while in this section we consider
the general abstract definition given in Defintion~\ref{def:opsystem}.

@@

As described beforehand in this section, the collection of Facts shall define
the functional and the physical architecture of an agent, but Assets and
security properties are defined as Facts, i.e. true. Given that understanding
why an Asset is considered as such by the user doesn't necessarily allow
us to better qualify an Asset, we consider the fact of ``being an asset'' as
a predicate of a component of a system.

@@

\subsection{Qualitative Evaluation of Agent Space in $\abftheory$}\label{sec:agentspace}
While a quantitative analysis reveals how many possible configurations of an
agent (i.e. a system) exist w.r.t. the $\abftheory$ theory (i.e. $54/125$ in RCC5), a
qualitative analysis of the different configurations describes
the configurations allowed by the $\abftheory$ theory, and how those configurations can be
categorized.  In Table~\ref{tab:5com} we provide the generic composition table
of RCC5 over 3 regions instantiated over $\abf$, which shows the whole state
space for a single agent. The color coding of the table represents the 
risk function as a risk matrix as we detail afterwards in this section.

As depicted in
Figure~\ref{fig:soundness}, the relation between Facts, and Assertions and Behavior
defines the soundness of the design (admissible configurations of Assertions,
Behaviors and, in turn, Beliefs) w.r.t. the specification (what the
specification mandates, such as by the nature of the physical/functional
architectural models). We categorize agents by first analyzing the relations
between each pair of Regions defining an agent (i.e. $\abf$), and then we
categorize the different agents as tuple of the three Regions.
For the sake of simplicity, soundness is opposed to non-soundness in the following, however,
with the RCC one should consider different ``degrees'' of non-soundness. For example,
in RCC5, if we consider EQ between two Regions as representing soundness, DR over the same Regions
represents non-soundness; while PP, PO, PPi represents the different degrees of non-soundness.
A similar argument can be done for completeness.

@@

\Paragraph{\Rcc{$\assertionRegion$}{$\behaviorRegion$} -- Collaboration}
In order to reason on the relation between assertions and behavior we first
need to consider that, by definition, assertions are defined as transfer of
information between two agents, e.g., $a$ and $b$.  Therefore, as depicted in
Figure~\ref{fig:system-agent}, an agent has two main categories of assertions,
input and output assertions.  Given an agent $a$ and a collection of asserted
predicates $\Phi$, the Input assertions are those received by $a$ from an agent
$s$ acting as a sender, $\rassert{s}{a}{\Phi}$; similarly, output assertions
are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. We shall consider
two pairs of regions\footnote{``I am not asserting, as Lotze did, that a relation between $X$ and $Y$
consists of a quality in $X$ and a quality in $Y$ -- a view which I regard as
quite indefensible. -- I assert that a relation $Z$ between $X$ and $Y$ involves
the existence in $X$ of the quality ``having the relation $Z$ to $Y$'' so that
a difference of relations always involves a difference in quality, and a
change of relations always involves a change of quality.'' --  Ellis J. McTaggart, The Unreality of Time\autocite{Mctaggart1908unreality}}: 
\begin{itemize}
	\item \Rcc{$\assertionRegion_{s\rightarrow a}$}{$\behaviorRegion$},
		where the relation between Input-assertions and behavior
		describes the soundness of the execution of the functional
		architecture w.r.t. input elicitation. With more details, the
		ideal specification of the functional architecture, along with
		the expected inputs, defines the functional behavior of an
		ideal system.  If all the inputs (Assertions) are correctly
		handled in the functional specification (behavior) the
		specification is complete.

@@

\Paragraph{\Rcc{$\assertionRegion$}{$\factRegion$} and \Rcc{$\behaviorRegion$}{$\factRegion$} -- Honesty and Competence}
While Assertions and Beliefs generated by the Behavior determines how a system should work, and the
relation between the two defines the quality (i.e. the completeness w.r.t the specification) of the agent, the relation of
Assertions and Behavior with Facts determines the quality with respect to the
nominal (specified) system.  Given that Facts defines what must be true in the
system, the relation of Assertions an Facts
determines the degree of quality between the information circulating in a
system (or within an agent) and reality\fixnote{mr}{assertions are reality
while facts are dogmatic assertions at specification level, i.e. requirements}.
Since the transfer of information through Assertions
generates Beliefs, a dishonest agent may circulate false information,
generating false Beliefs.  We note that it is often implied that the intention behind
circulating false information discerns a dishonest and an incompetent agent,
however, we consider honesty related to sharing truths\footnote{``[\ldots] truth
exists only in the good man, but the true in the bad man as well; for it is
possible for the bad man to utter something true.''
Sextus Empiricus, Outlines of Pyrrhonism, II-83\autocite{Empiricus1990Pyrrhonism}}.  The
relation between Beliefs and Facts determines the competence (on the subjects
defined by the Facts) of an agent (i.e. the more competent an agent is, the
more likely a Belief of that agent is true).
